### 2021-03-15 final-report.latex/main.tex
add discretized gradient flow

@@

Following \cite{bietenholz2018}, we can define the gradient flow in this model via the differential equation
\begin{equation}
    \label{eq:nsm_gradflow}
    \partial_\tau \vec\rho (\tau,x) = P(\tau,x) \nabla^2 \vec\rho(\tau,x),
\end{equation}
where $\nabla^2$ is the Laplacian operator in Euclidean space and $P(\tau,x)$ is a $N \times N$ matrix defined as
\begin{equation*}
    P(\tau, x) = 1 - \vec\rho(\tau,x) \vec\rho(\tau,x)^T.
\end{equation*}
We solve this equation numerically using the boundary condition $\vec\rho(0,x) = \phi(x)$.

In order to use this theory on the lattice, we discretize the action. We replace the derivatives with a simple difference and impose periodic boundary conditions on the system.

@@

Though the Metropolis algorithm will slowly find the absolute minimum of the theory, the presence of local minima can greatly prolong the convergence. Both the $\phi^4$ model and the nonlinear $\sigma$ model feature ``kinetic'' terms with gradients of $\phi$. Therefore, the presence of large similarly-valued regions in the lattice can lead to a local minimum. The Wolff algorithm helps reduce the presence of these clusters through two steps: identifying a cluster and flipping it along some arbitrary vector. In the case of $\phi^4$ theory, this flipping takes the form of a simple sign change. In the non-linear sigma model we choose a random unit vector $\vec r$ and consider the projection of the field on this vector. When the cluster flips, each site is flipped along this direction. \footnote{citation needed}. To identify the cluster, the algorithm uses a recursive algorithm defined by the probability of adding a new site. In the non-linear $\sigma$ model, this value is derived as 
\begin{equation}
    P_{add} = 1-e^{- \beta \vec\phi_a\vec\phi_b}.
\end{equation}
 % TODO: derive this

@@

>>>>

\newcommand{\e}{\vec e}

@@

Following \cite{bietenholz2018}, we can define the gradient flow in this model via the differential equation
\begin{equation}
    \label{eq:nsm_gradflow}
    \partial_\tau \vec\rho (\tau,x) = \left( 1 - \vec\rho(\tau,x) \vec\rho(\tau,x)^T \right) \nabla^2 \vec\rho(\tau,x),
\end{equation}
where $\nabla^2$ is the Laplacian operator in Euclidean space. We solve this equation numerically using the boundary condition $\vec\rho(0,x) = \phi(x)$.

In order to use this theory on the lattice, we discretize the action. We replace the derivatives with a simple difference and impose periodic boundary conditions on the system. For a more in-depth discussion, see Section~\ref{sec:grad_flow}.

@@

Though the Metropolis algorithm will slowly find the absolute minimum of the theory, the presence of local minima can greatly prolong the convergence. Both the $\phi^4$ model and the nonlinear $\sigma$ model feature ``kinetic'' terms with gradients of $\phi$. Therefore, the presence of large similarly-valued regions in the lattice can lead to a local minimum. The Wolff algorithm helps reduce the presence of these clusters through two steps: identifying a cluster and flipping it along some arbitrary vector. In the case of $\phi^4$ theory, this flipping takes the form of a simple sign change. In the non-linear sigma model we choose a random unit vector $\vec r$ and consider the projection of the field on this vector. When the cluster flips, each site is flipped along this direction. \footnote{citation needed}. To identify the cluster, the algorithm uses a recursive algorithm defined by the probability of adding a new site. In the non-linear $\sigma$ model, this value is derived as 
\begin{equation}
    P_{add} = 1-e^{- \beta \vec e_a\vec e_b}.
\end{equation}
 % TODO: derive this

@@

\section{Gradient Flow}
\label{sec:gradflow_disc}
On the lattice, the definition of the gradient flow (Eq.~\ref{eq:nsm_gradflow}) becomes 
\begin{equation}
    \label{eq:nsm_gradflow_disc}
    \partial_\tau \e (\tau)_x = \left( 1 - \e(\tau)_x \e(\tau)_x^T\right) \nabla^2 \e(\tau)_x,
\end{equation}
 where the Laplace operator $\nabla^2$ is defined as
\begin{equation*}
    \nabla^2 \e(t)_x = \sum^d_{\mu=1} \left( \e(t)_{x+\hat \mu} - \e(t)_{x-\hat \mu}\right) - 2 d \e(t)_x.
\end{equation*}

In order to calculate the gradient flow, we numerically solve the differential equation using a fourth-order Runga-Kutta approximation. We use an adaptive step size algorithm to dramatically increase efficiency.
